---
title: "ST2_Analysis"
author: "Gabriel Rojas"
date: "2025-02-24"
output:
  word_document: default
  html_document: default
---

```{r packages, include=FALSE}
library(gtsummary)
library(readxl)
library(dplyr)
library(stringr)
library(ggplot2)
library(plotly)
library(tidyr)
library(openxlsx)
library(readr)
library(gamlss)
library(gamlss2)
library(lubridate)
library(ggpubr)
library(stats)
library(patchwork)

source("eval_MixDist.R")
source("eval_ExtractMeta.R")
source("eval_Analyse.R")
```

```{r setup and settings, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi = 300,
                      echo = F, 
                      cache = T,
                      message = F, 
                      warning = F,
                      fig.height = 5,
                      fig.width = 6,
                      fig.align = 'center')

Perc_bins <- c(seq(0,100),0.5, 2.5,97.5,99.5)
Perc_bins <- Perc_bins[order(Perc_bins)]
CO2_bins <- seq(300, 5000, length.out = 150)
CO2_midbins <- CO2_bins[-length(CO2_bins)] + diff(CO2_bins)/2
PM25_bins <- seq(0, 600, length.out = 200)
T_bins <- seq(-20, 40, length.out = 61)
RH_bins <- seq(0, 100, length.out = 101)
HCHO_bins <- seq(0, 100, length.out = 101)
TVOC_bins <- seq(0, 3000, length.out = 151)
VOC_bins <- seq(0, 3000, length.out = 151)
Radon_bins <- seq(0, 3000, length.out = 151)
```

# read and filter data
```{r read data, include=FALSE}

stats <- readRDS("../tmp/stats.Rds")

filter_combi <- tibble(
  room0 = c("BED", "LIV", "KIT", "LAU", "HOO", "CCO", "EHA", "ETA"),
  tod  = c("23-07", "07-23", "07-23", "07-23", "07-23", "all", "all", "all"))

df <- stats %>%
  filter(!home=="General") %>% #avoid the "General" homes e.g. in Lodenareal for central mech system
  semi_join(filter_combi, by = c("room0", "tod"))  # Keep only matching rows
unique(df$study)
```

# prepare unaggregated data for summary grouped by room
```{r prepare unaggregated data for summary by room, message=TRUE}
df_pol <- df %>%
  filter(variable %in% c("T","RH","CO2","PM25","PM1","HCHO","TVOC","VOC","Radon")) %>%
  pivot_wider(id_cols = c("user","study","home","room","year","month","tod","room0","vent_type","country"),
              names_from = variable, values_from = Mean, # make sure to also change in df_1pt
              values_fn = list(Mean = mean)) %>%
  mutate(TVOC = coalesce(VOC, TVOC)) %>%
  select(-VOC)

# alternative values to extract !! nake sure to also change for df_1pt
# # median
# df_pol <- df %>%
#   unnest(perc) %>%
#   filter(variable %in% c("T","RH","CO2","PM25","PM1","HCHO","TVOC","VOC","Radon")) %>%
#   pivot_wider(id_cols = c("user","study","home","room","year","month","tod","room0","vent_type","country"),
#               names_from = variable, values_from = p50,
#               values_fn = list(p50 = mean)) %>%
#   mutate(TVOC = coalesce(VOC, TVOC)) %>%
#   select(-VOC)
# 
# # p95
# df_pol <- df %>%
#   unnest(perc) %>%
#   filter(variable %in% c("T","RH","CO2","PM25","PM1","HCHO","TVOC","VOC","Radon")) %>%
#   pivot_wider(id_cols = c("user","study","home","room","year","month","tod","room0","vent_type","country"),
#               names_from = variable, values_from = p95,
#               values_fn = list(p95 = mean)) %>%
#   mutate(TVOC = coalesce(VOC, TVOC)) %>%
#   select(-VOC)

```

# Summary table unaggregated data grouped by room
```{r summary table unaggregated data by room}
# Create summary table for with Mean values for BED
tbl_bed <- tbl_summary(df_pol[df_pol$room0=="BED", c("T", "RH", "CO2", "PM25", "HCHO", "TVOC", "Radon","vent_type")],
                      statistic = list(
                        all_continuous() ~ "{median}  \n({min}, {p5}, {p25}, {p75}, {p95}, {max}),  \n N={N_nonmiss}",
                        T ~ "{mean} ({sd}),  \n N={N_nonmiss}",
                        RH ~ "{mean} ({sd}),  \n N={N_nonmiss}"),
                      digits = list(
                        all_continuous() ~ 0,
                        T ~ 1),
                      by = vent_type, 
                      label = list(
                        T ~ "Temp. [°C)]",
                        RH ~ "RH [%]",
                        CO2 ~ "CO2 [ppm]",
                        PM25 ~ "PM2.5 [µg/m³]",
                        HCHO ~ "HCHO [µg/m³]",
                        TVOC ~ "VOC signal [µg/m³]",
                        Radon ~ "Radon [Bq/m³]"),
                      missing = "no") %>%
  add_overall(., last = FALSE) %>%
  modify_header(., all_stat_cols() ~ "**{level}**")
  #add_p(., pvalue_fun = ~ style_pvalue(.x, digits = 2)) %>%
  #modify_spanning_header(., c("stat_1", "stat_2", "stat_3", "stat_4") ~ "**Ventilation type**")

# Create summary table for with Mean values for LIV
tbl_liv <- tbl_summary(df_pol[df_pol$room0=="LIV", c("T", "RH", "CO2", "PM25", "HCHO", "TVOC", "Radon", "vent_type")],
                      statistic = list(
                        all_continuous() ~ "{median}  \n({min}, {p5}, {p25}, {p75}, {p95}, {max}),  \n N={N_nonmiss}",
                        T ~ "{mean} ({sd}),  \n N={N_nonmiss}",
                        RH ~ "{mean} ({sd}),  \n N={N_nonmiss}"),
                      digits = list(
                        all_continuous() ~ 0,
                        T ~ 1),
                      by = vent_type, 
                      label = list(
                        T ~ "Temp. [°C)]",
                        RH ~ "RH [%]",
                        CO2 ~ "CO2 [ppm]",
                        PM25 ~ "PM2.5 [µg/m³]",
                        HCHO ~ "HCHO [µg/m³]",
                        TVOC ~ "VOC signal [µg/m³]",
                        Radon ~ "Radon [Bq/m³]"),
                      missing = "no") %>%
  add_overall(., last = FALSE) %>%
  modify_header(., all_stat_cols() ~ "**{level}**")
  #add_p(., pvalue_fun = ~ style_pvalue(.x, digits = 2)) %>%
  #modify_spanning_header(., c("stat_1", "stat_2", "stat_3", "stat_4") ~ "**Ventilation type**")

# Create summary table for with Mean values for KIT
tbl_kit <- tbl_summary(df_pol[df_pol$room0=="KIT", c("T", "RH", "CO2", "PM25", "TVOC","Radon","vent_type")],
                      statistic = list(
                        all_continuous() ~ "{median}  \n({min}, {p5}, {p25}, {p75}, {p95}, {max}),  \n N={N_nonmiss}",
                        T ~ "{mean} ({sd}),  \n N={N_nonmiss}",
                        RH ~ "{mean} ({sd}),  \n N={N_nonmiss}"),
                      digits = list(
                        all_continuous() ~ 0,
                        T ~ 1),
                      by = vent_type, 
                      label = list(
                        T ~ "Temp. [°C)]",
                        RH ~ "RH [%]",
                        CO2 ~ "CO2 [ppm]",
                        PM25 ~ "PM2.5 [µg/m³]",
                        TVOC ~ "VOC signal [µg/m³]",
                        Radon ~ "Radon [Bq/m³]"),
                      missing = "no") %>%
  add_overall(., last = FALSE) %>%
  modify_header(., all_stat_cols() ~ "**{level}**")
  #add_p(., pvalue_fun = ~ style_pvalue(.x, digits = 2)) %>%
  #modify_spanning_header(., c("stat_1", "stat_2", "stat_3", "stat_4") ~ "**Ventilation type**")

# Create summary table for with Mean values for LAU (living area undefined) and HOO home office
tbl_lau <- tbl_summary(df_pol[df_pol$room0=="LAU" | df_pol$room0=="HOO", c("T", "RH", "CO2", "PM25", "HCHO", "TVOC","vent_type")],
                      statistic = list(
                        all_continuous() ~ "{median}  \n({min}, {p5}, {p25}, {p75}, {p95}, {max}),  \n N={N_nonmiss}",
                        T ~ "{mean} ({sd}),  \n N={N_nonmiss}",
                        RH ~ "{mean} ({sd}),  \n N={N_nonmiss}"),
                      digits = list(
                        all_continuous() ~ 0,
                        T ~ 1),
                      by = vent_type, 
                      label = list(
                        T ~ "Temp. [°C)]",
                        RH ~ "RH [%]",
                        CO2 ~ "CO2 [ppm]",
                        PM25 ~ "PM2.5 [µg/m³]",
                        HCHO ~ "HCHO [µg/m³]",
                        TVOC ~ "VOC signal [µg/m³]"),
                      missing = "no") %>%
  add_overall(., last = FALSE) %>%
  modify_header(., all_stat_cols() ~ "**{level}**")
  #add_p(., pvalue_fun = ~ style_pvalue(.x, digits = 2)) %>%
  #modify_spanning_header(., c("stat_1", "stat_2", "stat_3", "stat_4") ~ "**Ventilation type**")

# Create summary table for with Mean values for CCO (central corridor) and EHA/ETA (exhaust/extract flatwise)
tbl_cco <- tbl_summary(df_pol[df_pol$room0=="CCO" | df_pol$room0=="EHA" | df_pol$room0=="ETA", c("T", "RH", "CO2", "vent_type")],
                      statistic = list(
                        all_continuous() ~ "{median}  \n({min}, {p5}, {p25}, {p75}, {p95}, {max}),  \n N={N_nonmiss}",
                        T ~ "{mean} ({sd}),  \n N={N_nonmiss}",
                        RH ~ "{mean} ({sd}),  \n N={N_nonmiss}"),
                      digits = list(
                        all_continuous() ~ 0,
                        T ~ 1),
                      by = vent_type, 
                      label = list(
                        T ~ "Temp. [°C)]",
                        RH ~ "RH [%]",
                        CO2 ~ "CO2 [ppm]"),
                      missing = "no") %>%
  add_overall(., last = FALSE) %>%
  modify_header(., all_stat_cols() ~ "**{level}**")
  #add_p(., pvalue_fun = ~ style_pvalue(.x, digits = 2)) %>%
  #modify_spanning_header(., c("stat_1", "stat_2", "stat_3", "stat_4") ~ "**Ventilation type**")

# Stack tables
tbl <- tbl_stack(list(tbl_bed, tbl_liv, tbl_kit, tbl_lau, tbl_cco),
                 group_header = c("Bedrooms (23-07)", "Living rooms (07-23)", "Kitchen (07-23)", 
                                  "Undefined living area incl. home-office room (07-23)",
                                  "Central all-home measurement, e.g. extract duct, central corridor (0-24)"))

# Format tables----
tbl <- as_gt(tbl)
tbl <- gt::tab_style(tbl, style = gt::cell_text(weight = "bold"), locations = gt::cells_row_groups(groups = everything()))
tbl <- gt::opt_row_striping(tbl, row_striping = T)
tbl <- gt::tab_options(tbl, row.striping.background_color = "#E8EDF4", column_labels.background.color = "#3E5A82")
#tbl <- gt::fmt_scientific(tbl)
knitr::knit_print(tbl)

```

# prepare unaggregated data for summary grouped by pollutant
```{r prepare unaggregated data for summary by pollutant, message=TRUE}

df_room <- df %>%
  filter(room0 %in% c("BED","LIV","KIT","LAU","HOO","CCO","EHA")) %>%
  pivot_wider(id_cols = c("user","study","home","room","year","month","tod","variable","vent_type","country"),
              names_from = room0, values_from = Mean,
              values_fn = list(Mean = mean)) %>%
  mutate(LAU = coalesce(LAU, HOO)) %>%
  select(-HOO) %>%
  mutate(CCO = coalesce(CCO, EHA)) %>%
  select(-EHA)

# combine VOC and TVOC (has do be done with care!)
message("Changing VOC to TVOC. The following studies are affected: ")
print(unique(df_room$study[df_room$variable=="VOC"]), max.levels = 0)
df_room <- df_room %>%
  mutate(variable = if_else(variable=="VOC", "TVOC", variable))

# for (pol in c("T","RH","CO2","PM25","HCHO","TVOC","VOC","Radon")){
#   message("Creating df_wide for: ", pol)
#   dfa_tempo <- df[df$variable==pol,] %>%
#     filter(room0 %in% c("BED","LIV","KIT","LAU","HOO","CCO","EHA")) %>%
#     pivot_wider(id_cols = c("user","study","home","room","year","month","tod","variable","vent_type","country"),
#                 names_from = room0, values_from = Mean,
#                 values_fn = list(Mean = mean)) %>%
#     mutate(LAU = if ("LAU" %in% names(.)) coalesce(LAU, HOO)) %>%
#     select(-any_of("HOO")) %>%
#     mutate(CCO = if ("CCO" %in% names(.)) coalesce(CCO, EHA)) %>%
#     select(-any_of("EHA"))
#   assign(paste0("Mean0_wide_",pol), dfa_tempo)
# }
# message("Combining all TVOC and VOC measurements from TS/low-cost sensors")
# Mean0_wide_TVOC <- Mean0_wide_TVOC %>%
#   full_join(.,Mean0_wide_VOC) %>%
#   mutate(variable="TVOC")

```

# Summary table unaggregated data grouped by pollutant
```{r summary table unaggregated data by pollutant}
tbl_rooms <- list(T=c("BED", "LIV", "KIT", "LAU", "CCO", "vent_type"),
                  RH=c("BED", "LIV", "KIT", "LAU", "CCO", "vent_type"),
                  CO2=c("BED", "LIV", "KIT", "LAU", "CCO", "vent_type"),
                  PM25=c("BED", "LIV", "KIT", "LAU", "vent_type"),
                  HCHO=c("BED", "LIV", "LAU", "vent_type"),
                  TVOC=c("BED", "LIV", "KIT", "LAU", "vent_type"),
                  Radon=c("BED", "LIV", "KIT", "vent_type"))
tbl_label <- list(T=list(BED ~ "Bedroom (23-07)", LIV ~ "Living room (07-23)", KIT ~ "Kitchen (07-23)",
                         LAU ~ "Undef. living area / home-office", CCO ~ "central corridor / extract (0-24)"),
                  RH=list(BED ~ "Bedroom (23-07)", LIV ~ "Living room (07-23)", KIT ~ "Kitchen (07-23)",
                          LAU ~ "Undef. living area / home-office (07-23)", CCO ~ "central corridor / extract (0-24)"),
                  CO2=list(BED ~ "Bedroom (23-07)", LIV ~ "Living room (07-23)", KIT ~ "Kitchen (07-23)",
                           LAU ~ "Undef. living area / home-office (07-23)", CCO ~ "central corridor / extract"),
                  PM25=list(BED ~ "Bedroom (23-07)", LIV ~ "Living room (07-23)", KIT ~ "Kitchen (07-23)",
                            LAU ~ "Undef. living area / home-office (07-23)"),
                  HCHO=list(BED ~ "Bedroom (23-07)", LIV ~ "Living room (07-23)", LAU ~ "Undef. living area / home-office (07-23)"),
                  TVOC=list(BED ~ "Bedroom (23-07)", LIV ~ "Living room (07-23)", KIT ~ "Kitchen (07-23)"),
                  Radon=list(BED ~ "Bedroom (23-07)", LIV ~ "Living room (07-23)", KIT ~ "Kitchen (07-23)"))
tbl_header <- list(T="Temperature [°C]", RH="Relative Humidiy [%]", CO2="CO2 [ppm]", PM25="PM2.5 [µg/m³]",
                                  HCHO="Formaldehyde [µg/m³]", TVOC="TVOC (incl. low cost sensor) [µg/m³]", Radon="Radon [Bq/m³]")

for (pol in c("T","RH","CO2","PM25","HCHO","TVOC","Radon")){
  message("Creating table for ", pol)
  #tbl_x <- tbl_summary(get(paste0("Mean0_wide_",pol))[, tbl_rooms[[pol]] ],
  tbl_x <- tbl_summary(df_room[df_room$variable==pol, tbl_rooms[[pol]] ],
                       statistic = all_continuous() ~ ifelse(pol=="T" | pol=="RH", "{mean} ({sd}),  \n N={N_nonmiss}",
                                                             "{median}  \n({min}, {p5}, {p25}, {p75}, {p95}, {max}),  \n N={N_nonmiss}"),
                       digits = all_continuous() ~ ifelse(pol=="T", 1, 0),
                        by = vent_type, 
                        label = tbl_label[[pol]],
                        missing = "no") %>%
    add_overall(., last = FALSE) %>%
    modify_header(., label = tbl_header[[pol]]) %>%
    modify_header(., all_stat_cols() ~ "**{level}**")
    #add_p(., pvalue_fun = ~ style_pvalue(.x, digits = 2)) %>%
    #modify_spanning_header(., c("stat_1", "stat_2", "stat_3", "stat_4") ~ "**Ventilation type**")
  assign(paste0("tbl_s_",pol), as_gt(tbl_x) %>%
           gt::tab_style(., style = gt::cell_text(weight = "bold"), locations = gt::cells_row_groups(groups = everything())) %>%
           gt::opt_row_striping(., row_striping = T) %>%
           gt::tab_options(., row.striping.background_color = "#E8EDF4", column_labels.background.color = "#3E5A82"))
  assign(paste0("tbl_",pol), tbl_x %>%
           modify_header(., label ~ ""))
  
}

# Stack tables
tbl <- tbl_stack(list(tbl_T, tbl_RH, tbl_CO2, tbl_PM25, tbl_HCHO, tbl_TVOC, tbl_Radon),
                 group_header = c("Temperature [°C]", "Relative Humidiy [%]", "CO2 [ppm]", "PM2.5 [µg/m³]",
                                  "Formaldehyde [µg/m³]", "TVOC (incl. low cost sensor) [µg/m³]", "Radon [Bq/m³]"))
# Format table----
tbl <- as_gt(tbl)
tbl <- gt::tab_style(tbl, style = gt::cell_text(weight = "bold"), locations = gt::cells_row_groups(groups = everything()))
tbl <- gt::opt_row_striping(tbl, row_striping = T)
tbl <- gt::tab_options(tbl, row.striping.background_color = "#E8EDF4", column_labels.background.color = "#3E5A82")
#tbl <- gt::fmt_scientific(tbl)
#knitr::knit_print(tbl)
knitr::knit_print(tbl_s_T)
knitr::knit_print(tbl_s_RH)
knitr::knit_print(tbl_s_CO2)
knitr::knit_print(tbl_s_PM25)
knitr::knit_print(tbl_s_HCHO)
knitr::knit_print(tbl_s_TVOC)
knitr::knit_print(tbl_s_Radon)
#list(knitr::knit_print(tbl), knitr::knit_print(tbl_s_T), knitr::knit_print(tbl_s_RH),knitr::knit_print(tbl_s_CO2), knitr::knit_print(tbl_s_PM25), knitr::knit_print(tbl_s_HCHO), knitr::knit_print(tbl_s_TVOC), knitr::knit_print(tbl_s_Radon))
```

# Summary plot unaggregated data
```{r plot }
n_T <- df_pol[df_pol$room0 %in% c("BED","LIV", "KIT") & !is.na(df_pol$T),] %>%
  group_by(room0,vent_type) %>%
  summarise(N_count = n(), R_count = length(unique(paste0(study,home,room))))

n_CO2 <- df_pol[df_pol$room0 %in% c("BED","LIV", "KIT") & !is.na(df_pol$CO2),] %>%
  group_by(room0,vent_type) %>%
  summarise(N_count = n(), R_count = length(unique(paste0(study,home,room))))

n_PM25 <- df_pol[df_pol$room0 %in% c("BED","LIV", "KIT") & !is.na(df_pol$PM25),] %>%
  group_by(room0,vent_type) %>%
  summarise(N_count = n(), R_count = length(unique(paste0(study,home,room))))

n_HCHO <- df_pol[df_pol$room0 %in% c("BED","LIV", "KIT") & !is.na(df_pol$HCHO),] %>%
  group_by(room0,vent_type) %>%
  summarise(N_count = n(), R_count = length(unique(paste0(study,home,room))))

p1 <- ggplot(df_pol[df_pol$room0 %in% c("BED","LIV", "KIT"),], aes(x = room0, y = T, fill = vent_type)) +
  #geom_violin(fill = "lightblue", color = "black") +
  geom_boxplot() +
  scale_x_discrete(labels = c("BED" = "Bedroom (23-07)", "LIV" = "Living Room (07-23)", "KIT" = "Kitchen (07-23)")) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 2, shape = 1, position = position_dodge(0.8)) + # Mittlere Werte
  geom_text(data = n_T, aes(x = room0, y = 12, label = paste0("N=",N_count)),
            size = 3, position = position_dodge(0.8)) +
  geom_text(data = n_T, aes(x = room0, y = 11, label = paste0("n=",R_count)),
            size = 3, position = position_dodge(0.8)) +
  geom_hline(yintercept = 30, linetype = "dashed", color = "black") +  # Cutoff-Grenze
  annotate("text", x = 0.5, y = 30, label = "outliers above not shown", hjust=0, vjust = -0.2, size = 3, color = "black") + # Hinweistext
  theme_minimal() +
  coord_cartesian(ylim=c(10,30)) +
  labs(x = "Room",
        y = expression("Mean Temperature [°C]"))
p1

p2 <- ggplot(df_pol[df_pol$room0 %in% c("BED","LIV", "KIT"),], aes(x = room0, y = CO2, fill = vent_type)) +
  geom_boxplot() +
  scale_x_discrete(labels = c("BED" = "Bedroom (23-07)", "LIV" = "Living Room (07-23)", "KIT" = "Kitchen (07-23)")) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 2, shape = 1, position = position_dodge(0.8)) + # Mittlere Werte
  geom_text(data = n_CO2, aes(x = room0, y = 175, label = paste0("N=",N_count)),
            size = 3, position = position_dodge(0.8)) +
  geom_text(data = n_CO2, aes(x = room0, y = 75, label = paste0("n=",R_count)), 
            size = 3, position = position_dodge(0.8)) +
  geom_hline(yintercept = 2000, linetype = "dashed", color = "black") +  # Cutoff-Grenze
  annotate("text", x = 0.5, y = 2000, label = "outliers above not shown", hjust=0, vjust = -0.2, size = 3, color = "black") + # Hinweistext
  theme_minimal() +
  coord_cartesian(ylim=c(0,2000)) +
  labs(x = "Room",
       y = expression("Mean CO"[2]~ "concentration [ppm]"))
p2

p3 <- ggplot(df_pol[df_pol$room0 %in% c("BED","LIV", "KIT"),], aes(x = room0, y = PM25, fill = vent_type)) +
  geom_boxplot() +
  scale_x_discrete(labels = c("BED" = "Bedroom (23-07)", "LIV" = "Living Room (07-23)", "KIT" = "Kitchen (07-23)")) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 2, shape = 1, position = position_dodge(0.8)) + # Mittlere Werte
  geom_text(data = n_PM25, aes(x = room0, y = 140, label = paste0("N=",N_count)),
            size = 3, position = position_dodge(0.8)) +
  geom_text(data = n_PM25, aes(x = room0, y = 130, label = paste0("n=",R_count)), 
            size = 3, position = position_dodge(0.8)) +
  geom_hline(yintercept = 150, linetype = "dashed", color = "black") +  # Cutoff-Grenze
  annotate("text", x = 0.5, y = 150, label = "outliers above not shown", hjust=0, vjust = -0.2, size = 3, color = "black") + # Hinweistext
  theme_minimal() +
  coord_cartesian(ylim=c(0,150)) +
  labs(x = "Room",
       y = expression("Mean PM"[25]~ "concentration [µg/m³]"))
p3

p4 <- ggplot(df_pol[df_pol$room0 %in% c("BED","LIV"),], aes(x = room0, y = HCHO, fill = vent_type)) +
  geom_boxplot() +
  scale_x_discrete(labels = c("BED" = "Bedroom (23-07)", "LIV" = "Living Room (07-23)")) +
  stat_summary(fun = mean, geom = "point", color = "black", size = 2, shape = 1, position = position_dodge(0.8)) + # Mittlere Werte
  geom_text(data = n_HCHO, aes(x = room0, y = 95, label = paste0("N=",N_count)),
            size = 3, position = position_dodge(0.8)) +
  geom_text(data = n_HCHO, aes(x = room0, y = 88, label = paste0("n=",R_count)), 
            size = 3, position = position_dodge(0.8)) +
  geom_hline(yintercept = 100, linetype = "dashed", color = "black") +  # Cutoff-Grenze
  annotate("text", x = 0.5, y = 100, label = "outliers above not shown", hjust=0, vjust = -0.2, size = 3, color = "black") + # Hinweistext
  theme_minimal() +
  coord_cartesian(ylim=c(0,100)) +
  labs(x = "Room",
       y = expression("Mean HCHO concentration [µg/m³]"))
p4

(p1 + p2) / (p3 + p4) +
plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom",  # Move legend below
    text = element_text(size = 10),
    plot.caption = element_text(size = 8, face = "italic") )
```

# prepare aggregated data for summary grouped by pollutant (aggregated to one data point per room)
```{r prepare aggregated TS-data for summary by pollutant, message=TRUE}

#dfa_room <- data.frame()
for (pol in c("T","RH","CO2","PM25","HCHO","TVOC","VOC","Radon")){
  message("Aggregating ", pol)
  Pol_bins <- get(paste0(pol,"_bins"))
  #assign(paste0("dfa_",pol), aggRoom(df[df$variable==pol,],Perc_bins,Pol_bins))
  dfa_xxx <- aggRoom(df[df$variable==pol,],Perc_bins,Pol_bins)
  
  dfa_xxx <- dfa_xxx %>%
    filter(room0 %in% c("BED","LIV","KIT","LAU","HOO","CCO","EHA")) %>%
    pivot_wider(id_cols = c("user","study","home","room","year","month","tod","variable","vent_type","country"),
                names_from = room0, values_from = Mean,
                values_fn = list(Mean = mean)) %>%
    mutate(LAU = if ("LAU" %in% names(.)) coalesce(LAU, HOO)) %>%
    select(-any_of("HOO")) %>%
    mutate(CCO = if ("CCO" %in% names(.)) coalesce(CCO, EHA)) %>%
    select(-any_of("EHA"))
  
  #dfa_room <- dfa_room %>%
  #  full_join(.,dfa_xxx)
  assign(paste0("dfa_room_",pol), dfa_xxx)
}
message("Changing VOC to TVOC. The following studies are affected: ")
print(unique(dfa_room_VOC$study), max.levels = 0)
dfa_room_TVOC <- dfa_room_TVOC %>%
  full_join(.,dfa_room_VOC) %>%
  mutate(variable="TVOC")

rm(dfa_room_VOC)

#dfa_PM <- aggRoom(df[df$variable=="PM25",],Perc_bins,PM_bins)
#unique(dfa_PM$study)
#dfa1 <- stats[stats$variable=="PM25" & stats$year =="all" & stats$month=="all" & stats$tod=="all",]
#unique(dfa1$study) #not all studies have the full aggregation all, all, all
```

# Summary table aggregated data by pollutant
```{r summary table aggregated TS-data by pollutant}
tbl_rooms <- list(T=c("BED", "LIV", "KIT", "LAU", "CCO", "vent_type"),
                  RH=c("BED", "LIV", "KIT", "LAU", "CCO", "vent_type"),
                  CO2=c("BED", "LIV", "KIT", "LAU", "CCO", "vent_type"),
                  PM25=c("BED", "LIV", "KIT", "LAU", "vent_type"),
                  HCHO=c("BED", "LIV", "LAU", "vent_type"),
                  TVOC=c("BED", "LIV", "KIT", "LAU", "vent_type"),
                  Radon=c("BED", "LIV", "KIT", "vent_type"))
tbl_label <- list(T=list(BED ~ "Bedroom (23-07)", LIV ~ "Living room", KIT ~ "Kitchen", LAU ~ "Undef. living area / home-office", CCO ~ "central corridor / extract"),
                  RH=list(BED ~ "Bedroom", LIV ~ "Living room", KIT ~ "Kitchen", LAU ~ "Undef. living area / home-office", CCO ~ "central corridor / extract"),
                  CO2=list(BED ~ "Bedroom", LIV ~ "Living room", KIT ~ "Kitchen", LAU ~ "Undef. living area / home-office", CCO ~ "central corridor / extract"),
                  PM25=list(BED ~ "Bedroom", LIV ~ "Living room", KIT ~ "Kitchen", LAU ~ "Undef. living area / home-office"),
                  HCHO=list(BED ~ "Bedroom", LIV ~ "Living room", LAU ~ "Undef. living area / home-office"),
                  TVOC=list(BED ~ "Bedroom", LIV ~ "Living room", KIT ~ "Kitchen"),
                  Radon=list(BED ~ "Bedroom", LIV ~ "Living room", KIT ~ "Kitchen"))
tbl_header <- list(T="Temperature [°C]", RH="Relative Humidiy [%]", CO2="CO2 [ppm]", PM25="PM2.5 [µg/m³]",
                                  HCHO="Formaldehyde [µg/m³]", TVOC="TVOC (incl. low cost sensor) [µg/m³]", Radon="Radon [Bq/m³]")

for (pol in c("T","RH","CO2","PM25","HCHO","TVOC","Radon")){
  message("Creating table for ", pol)
  tbl_x <- tbl_summary(get(paste0("dfa_room_",pol))[, tbl_rooms[[pol]] ],
                       statistic = all_continuous() ~ ifelse(pol=="T" | pol=="RH", "{mean} ({sd}),  \n N={N_nonmiss}",
                                                             "{median}  \n({min}, {p5}, {p25}, {p75}, {p95}, {max}),  \n N={N_nonmiss}"),
                       digits = all_continuous() ~ ifelse(pol=="T", 1, 0),
                        by = vent_type, 
                        label = tbl_label[[pol]],
                        missing = "no") %>%
    add_overall(., last = FALSE) %>%
    modify_header(., label = tbl_header[[pol]]) %>%
    modify_header(., all_stat_cols() ~ "**{level}**")
    #add_p(., pvalue_fun = ~ style_pvalue(.x, digits = 2)) %>%
    #modify_spanning_header(., c("stat_1", "stat_2", "stat_3", "stat_4") ~ "**Ventilation type**")
  assign(paste0("tbl_s_",pol), as_gt(tbl_x) %>%
           gt::tab_style(., style = gt::cell_text(weight = "bold"), locations = gt::cells_row_groups(groups = everything())) %>%
           gt::opt_row_striping(., row_striping = T) %>%
           gt::tab_options(., row.striping.background_color = "#E8EDF4", column_labels.background.color = "#3E5A82"))
  assign(paste0("tbl_",pol), tbl_x %>%
           modify_header(., label ~ ""))
}

# Stack tables
tbl <- tbl_stack(list(tbl_T, tbl_RH, tbl_CO2, tbl_PM25, tbl_HCHO, tbl_TVOC, tbl_Radon),
                 group_header = c("Temperature [°C]", "Relative Humidiy [%]", "CO2 [ppm]", "PM2.5 [µg/m³]",
                                  "Formaldehyde [µg/m³]", "TVOC (incl. low cost sensor) [µg/m³]", "Radon [Bq/m³]"))

# Format tables----
tbl <- as_gt(tbl)
tbl <- gt::tab_style(tbl, style = gt::cell_text(weight = "bold"), locations = gt::cells_row_groups(groups = everything()))
tbl <- gt::opt_row_striping(tbl, row_striping = T)
tbl <- gt::tab_options(tbl, row.striping.background_color = "#E8EDF4", column_labels.background.color = "#3E5A82")
#tbl <- gt::fmt_scientific(tbl)
knitr::knit_print(tbl_s_T)
knitr::knit_print(tbl_s_RH)
knitr::knit_print(tbl_s_CO2)
knitr::knit_print(tbl_s_PM25)
knitr::knit_print(tbl_s_HCHO)
knitr::knit_print(tbl_s_TVOC)
knitr::knit_print(tbl_s_Radon)
#list(knitr::knit_print(tbl), knitr::knit_print(tbl_s_T), knitr::knit_print(tbl_s_RH),knitr::knit_print(tbl_s_CO2), knitr::knit_print(tbl_s_PM25), knitr::knit_print(tbl_s_HCHO), knitr::knit_print(tbl_s_TVOC), knitr::knit_print(tbl_s_Radon))
```

# Combining with non time series (TS) data, e.g. passiv samplers (with only N=1, e.g. in Lueftung3 and IECH)
```{r combining with single point data, message=TRUE}
df_1pt <- stats %>%
  filter(study %in% c("IECH","Lueftung3")) %>% #the two studies where single measurements have been documented: IECH and Lueftung3 
  filter(N==1 | is.na(N)) %>%
  mutate(year=as.numeric(year),
         month=as.numeric(as.character(month)) )

# make sure they have same pollutant name
df_1pt$variable[df_1pt$study=="IECH" & df_1pt$variable=="Other12"] <- "Der_p1"
df_1pt$variable[df_1pt$study=="Lueftung3" & df_1pt$variable=="Mites (Der p1)"] <- "Der_p1"
df_1pt$variable[df_1pt$study=="IECH" & df_1pt$variable=="Other13"] <- "Der_f1"
df_1pt$variable[df_1pt$study=="Lueftung3" & df_1pt$variable=="Mites (Der f1)"] <- "Der_f1"
df_1pt$variable[df_1pt$study=="Lueftung3" & df_1pt$variable=="RAD"] <- "Radon"

df_1pt <- df_1pt %>%
  filter(room0 %in% c("BED","LIV","KIT","LAU")) %>%
  pivot_wider(id_cols = c("user","study","home","room","year","month","tod","variable","vent_type","country"),
              names_from = room0, values_from = Mean,
              values_fn = list(Mean = mean))


for (pol in c("Der_f1", "Der_p1", "HCHO", "TVOC", "Fungi", "Radon")){
  message("Merging 1-point data: ", pol)
  if (exists(paste0("dfa_room_", pol))){
    dfa_xxx <- df_1pt[df_1pt$variable==pol,] %>%
      full_join(.,get(paste0("dfa_room_", pol))) #%>%
      #filter(!if_all(c(LIV, BED, KIT), is.na))
    assign(paste0("dfa1_room_",pol),dfa_xxx)
  } else {
    assign(paste0("dfa1_room_",pol),df_1pt[df_1pt$variable==pol,])
  }
}

```
# Summary table all data (agg + 1 pt) by pollutant
```{r summary table all data by pollutant}
tbl_rooms <- list(T=c("BED", "LIV", "KIT", "LAU", "CCO", "vent_type"),
                  RH=c("BED", "LIV", "KIT", "LAU", "CCO", "vent_type"),
                  CO2=c("BED", "LIV", "KIT", "LAU", "CCO", "vent_type"),
                  PM25=c("BED", "LIV", "KIT", "LAU", "vent_type"),
                  HCHO=c("BED", "LIV", "LAU", "vent_type"),
                  TVOC=c("BED", "LIV", "KIT", "LAU", "vent_type"),
                  Radon=c("BED", "LIV", "KIT", "vent_type"))
tbl_label <- list(T=list(BED ~ "Bedroom (23-07)", LIV ~ "Living room", KIT ~ "Kitchen", LAU ~ "Undef. living area / home-office", CCO ~ "central corridor / extract"),
                  RH=list(BED ~ "Bedroom", LIV ~ "Living room", KIT ~ "Kitchen", LAU ~ "Undef. living area / home-office", CCO ~ "central corridor / extract"),
                  CO2=list(BED ~ "Bedroom", LIV ~ "Living room", KIT ~ "Kitchen", LAU ~ "Undef. living area / home-office", CCO ~ "central corridor / extract"),
                  PM25=list(BED ~ "Bedroom", LIV ~ "Living room", KIT ~ "Kitchen", LAU ~ "Undef. living area / home-office"),
                  HCHO=list(BED ~ "Bedroom", LIV ~ "Living room", LAU ~ "Undef. living area / home-office"),
                  TVOC=list(BED ~ "Bedroom", LIV ~ "Living room", KIT ~ "Kitchen"),
                  Radon=list(BED ~ "Bedroom", LIV ~ "Living room", KIT ~ "Kitchen"))
tbl_header <- list(T="Temperature [°C]", RH="Relative Humidiy [%]", CO2="CO2 [ppm]", PM25="PM2.5 [µg/m³]",
                                  HCHO="Formaldehyde [µg/m³]", TVOC="TVOC (incl. low cost sensor) [µg/m³]", Radon="Radon [Bq/m³]")

for (pol in c("T","RH","CO2","PM25","HCHO","TVOC","Radon")){
  message("Creating table for ", pol)
  tbl_x <- tbl_summary(get(paste0("dfa_room_",pol))[, tbl_rooms[[pol]] ],
                       statistic = all_continuous() ~ ifelse(pol=="T" | pol=="RH", "{mean} ({sd}),  \n N={N_nonmiss}",
                                                             "{median}  \n({min}, {p5}, {p25}, {p75}, {p95}, {max}),  \n N={N_nonmiss}"),
                       digits = all_continuous() ~ ifelse(pol=="T", 1, 0),
                        by = vent_type, 
                        label = tbl_label[[pol]],
                        missing = "no") %>%
    add_overall(., last = FALSE) %>%
    modify_header(., label = tbl_header[[pol]]) %>%
    modify_header(., all_stat_cols() ~ "**{level}**")
    #add_p(., pvalue_fun = ~ style_pvalue(.x, digits = 2)) %>%
    #modify_spanning_header(., c("stat_1", "stat_2", "stat_3", "stat_4") ~ "**Ventilation type**")
  assign(paste0("tbl_s_",pol), as_gt(tbl_x) %>%
           gt::tab_style(., style = gt::cell_text(weight = "bold"), locations = gt::cells_row_groups(groups = everything())) %>%
           gt::opt_row_striping(., row_striping = T) %>%
           gt::tab_options(., row.striping.background_color = "#E8EDF4", column_labels.background.color = "#3E5A82"))
  assign(paste0("tbl_",pol), tbl_x %>%
           modify_header(., label ~ ""))
}

# Stack tables
tbl <- tbl_stack(list(tbl_T, tbl_RH, tbl_CO2, tbl_PM25, tbl_HCHO, tbl_TVOC, tbl_Radon),
                 group_header = c("Temperature [°C]", "Relative Humidiy [%]", "CO2 [ppm]", "PM2.5 [µg/m³]",
                                  "Formaldehyde [µg/m³]", "TVOC (incl. low cost sensor) [µg/m³]", "Radon [Bq/m³]"))

# Format tables----
tbl <- as_gt(tbl)
tbl <- gt::tab_style(tbl, style = gt::cell_text(weight = "bold"), locations = gt::cells_row_groups(groups = everything()))
tbl <- gt::opt_row_striping(tbl, row_striping = T)
tbl <- gt::tab_options(tbl, row.striping.background_color = "#E8EDF4", column_labels.background.color = "#3E5A82")
#tbl <- gt::fmt_scientific(tbl)
knitr::knit_print(tbl)
knitr::knit_print(tbl_s_T)
knitr::knit_print(tbl_s_RH)
knitr::knit_print(tbl_s_CO2)
knitr::knit_print(tbl_s_PM25)
knitr::knit_print(tbl_s_HCHO)
knitr::knit_print(tbl_s_TVOC)
knitr::knit_print(tbl_s_Radon)
#list(knitr::knit_print(tbl), knitr::knit_print(tbl_s_T), knitr::knit_print(tbl_s_RH),knitr::knit_print(tbl_s_CO2), knitr::knit_print(tbl_s_PM25), knitr::knit_print(tbl_s_HCHO), knitr::knit_print(tbl_s_TVOC), knitr::knit_print(tbl_s_Radon))
```


<!-- # Check distribution -->

<!-- ```{r check distribution of variables, eval=FALSE} -->

<!-- # Distribution of HCHO -->

<!-- length(df$Mean) -->

<!-- sum(is.na(df$Mean)) -->

<!-- hist(df$Mean) -->

<!-- ggqqplot(df$Mean) -->

<!-- shapiro.test(df$Mean[1:5000]) -->

<!-- df_test <- df[df$Mean != 0,] -->

<!-- head(log(df_test$Mean)) -->

<!-- sum(is.na(log(df$Mean))) -->

<!-- hist(log(df_test$Mean)) -->

<!-- ggqqplot(log(df$Mean)) -->

<!-- shapiro.test(log(df_test$Mean[1:5000])) -->

<!-- ``` -->

<!-- # box plots -->

<!-- ```{r check PM25 concentrations across studies, eval=FALSE} -->

<!-- ggplot(df, aes(x=room0, y=Mean, fill=country)) +  -->

<!--   geom_boxplot() + -->

<!--   xlab("") + -->

<!--   ylab("TVOC [µg/m3]") + -->

<!--   labs(fill = "") #+ -->

<!--   #ylim(0, 200) + -->

<!--   #labs(tag = "(a)") + -->

<!--   #scale_fill_manual(values = c("#C9D4E5", "#6989B7")) + -->

<!--   # theme(panel.background = element_blank(), -->

<!--   #       panel.border = element_blank(), -->

<!--   #       axis.line = element_line(size = 0.5, linetype = "solid", colour = "black"), -->

<!--   #       axis.text.x = element_text(size = 10, face = "plain"), -->

<!--   #       axis.text.y = element_text(size = 10, face = "plain"), -->

<!--   #       axis.title.x = element_text(size = 10, face = "plain"), -->

<!--   #       axis.title.y = element_text(size = 10, face = "plain"), -->

<!--   #       legend.position = c(0.2, 0.9)) -->

<!-- ``` -->
